Runtime Feedback in a Meta-Tracing JIT for Efficient Dynamic Languages

### 2011-12-19 talk/icooolps2011/talk/talk.tex
this was done by the list-strategies branch
typo in talk slides

\begin{frame}
  \frametitle{A Tracing JIT}
  \includegraphics[scale=0.5]{figures/trace01.pdf}
\end{frame}

\begin{frame}
  \frametitle{An Interpreter}
  \includegraphics[scale=0.5]{figures/trace02.pdf}
\end{frame}

@@

\begin{frame}
  \frametitle{Example: Attribute Reads in Python}
  What happens when an attribute \texttt{x.m} is read? (simplified)
  \begin{itemize}
      \item check for \texttt{x.\_\_getattribute\_\_}, if there, call it
      \pause
      \item look for the attribute in the object's dictionary, if it's there, return it
      \pause
      \item walk up the MRO and look in each class' dictionary for the attribute
      \pause
      \item if the attribute is found, call its \texttt{\_\_get\_\_} attribute and return the result
      \pause
      \item if the attribute is not found, look for \texttt{x.\_\_getattr\_\_}, if there, call it
      \pause
      \item raise an \texttt{AttributeError}
  \end{itemize}
  \pause
  all this is one bytecode
\end{frame}

@@

\begin{frame}
  \frametitle{Meta-Tracing JITs}
  \begin{block}{Advantages:}
    \begin{itemize}
        \item semantics is always like that of the interpreter
        \item trace fully contains language semantics
        \item meta-tracers can be reused for various interpreters
    \end{itemize}
  \end{block}
  \pause
  a few meta-tracing systems have been built:
  \begin{itemize}
      \item Sullivan et.al. describe a meta-tracer using the Dynamo RIO system
      \item Yermolovich et.al. run a Lua implementation on top of a tracing JS implementation
      \item SPUR is a tracing JIT for CLR bytecodes, which is used to speed up a JS implementation in C\#
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Example: Instances with Maps}
  \includegraphics[scale=0.5]{figures/map01.pdf}
\end{frame}

\begin{frame}
  \frametitle{Example: Instances with Maps}
  \includegraphics[scale=0.5]{figures/map02.pdf}
\end{frame}

@@

\begin{frame}[plain,containsverbatim]
\frametitle{Trace for code \texttt{inst.a + inst.b}}
\begin{lstlisting}[mathescape,escapechar=|,basicstyle=\ttfamily]]
# $inst_1$.getfield("a")
$map_1$ = $inst_1$.map
$index_1$ = Map.getindex($map_1$, "a")
guard($index_1$ != -1)
$storage_1$ = $inst_1$.storage
$result_1$ = $storage_1$[$index_1$]
|\pause|
# $inst_1$.getfield("b")
$map_2$ = $inst_1$.map
$index_2$ = Map.getindex($map_2$, "b")
guard($index_2$ != -1)
$storage_2$ = $inst_1$.storage
$result_2$ = $storage_2$[$index_2$]

@@

    \PY{n+nd}{@elidable}
    \PY{k}{def} \PY{n+nf}{add\PYZus{}attribute}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{name}\PY{p}{)}\PY{p}{:}
        \PY{o}{.}\PY{o}{.}\PY{o}{.}

@@

\begin{frame}[containsverbatim,plain]
  \frametitle{Trace with Hints for Code \texttt{inst.a + inst.b}}
\begin{lstlisting}[mathescape,escapechar=|,basicstyle=\ttfamily]]
# $inst_1$.getfield("a")
$map_1$ = $inst_1$.map
guard($map_1$ == 0xb74af4a8)
|{\color{gray}$index_1$ = Map.getindex($map_1$, "a")}|
|{\color{gray}guard($index_1$ != -1)}|
$storage_1$ = $inst_1$.storage
$result_1$ = $storage_1$[$index_1$}]

>>>>

\begin{frame}
  \frametitle{An Interpreter}
  \includegraphics[scale=0.5]{figures/trace01.pdf}
\end{frame}

\begin{frame}
  \frametitle{A Tracing JIT}
  \includegraphics[scale=0.5]{figures/trace02.pdf}
\end{frame}

@@

\begin{frame}
  \frametitle{Example: Attribute Reads in Python}
  What happens when an attribute \texttt{x.m} is read? (simplified)
  \pause
  \begin{itemize}
      \item check for \texttt{x.\_\_getattribute\_\_}, if there, call it
      \pause
      \item look for the attribute in the object's dictionary, if it's there, return it
      \pause
      \item walk up the MRO and look in each class' dictionary for the attribute
      \pause
      \item if the attribute is found, call its \texttt{\_\_get\_\_} attribute and return the result
      \pause
      \item if the attribute is not found, look for \texttt{x.\_\_getattr\_\_}, if there, call it
      \pause
      \item raise an \texttt{AttributeError}
  \end{itemize}
  \pause
  all this is one bytecode
\end{frame}

@@

\begin{frame}
  \frametitle{Meta-Tracing JITs}
  \begin{block}{Advantages:}
    \begin{itemize}
        \item semantics are always like that of the interpreter
        \item trace fully contains language semantics
        \item meta-tracers can be reused for various interpreters
    \end{itemize}
  \end{block}
  \pause
  a few meta-tracing systems have been built:
  \begin{itemize}
      \item Sullivan et.al. describe a meta-tracer using the Dynamo RIO system
      \item Yermolovich et.al. run a Lua implementation on top of a tracing JS implementation
      \item SPUR is a tracing JIT for CLR bytecodes, which is used to speed up a JS implementation in C\#
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Example: Instances with Maps}
  \includegraphics[scale=0.7]{figures/map01.pdf}
\end{frame}

\begin{frame}
  \frametitle{Example: Instances with Maps}
  \includegraphics[scale=0.7]{figures/map02.pdf}
\end{frame}

\begin{frame}
  \frametitle{Example: Instances with Maps}
  \includegraphics[scale=0.7]{figures/map03.pdf}
\end{frame}

@@

\begin{frame}[plain,containsverbatim]
\frametitle{Trace for Code \texttt{inst.a + inst.b}}
\begin{lstlisting}[mathescape,escapechar=|,basicstyle=\ttfamily]]
# $inst_1$.getfield("a")
$map_1$ = $inst_1$.map
$index_1$ = Map.getindex($map_1$, "a")
guard($index_1$ != -1)
$storage_1$ = $inst_1$.storage
$result_1$ = $storage_1$[$index_1$]
|\pause|
# $inst_1$.getfield("b")
$map_2$ = $inst_1$.map
$index_2$ = Map.getindex($map_2$, "b")
guard($index_2$ != -1)
$storage_2$ = $inst_1$.storage
$result_2$ = $storage_2$[$index_2$]

@@

@@

\begin{frame}[containsverbatim,plain]
  \frametitle{Trace with Hints for Code \texttt{inst.a + inst.b}}
\begin{lstlisting}[mathescape,escapechar=|,basicstyle=\ttfamily]]
# $inst_1$.getfield("a")
$map_1$ = $inst_1$.map
guard($map_1$ == 0xb74af4a8)
|{\color{gray}$index_1$ = Map.getindex($map_1$, "a")}|
|{\color{gray}guard($index_1$ != -1)}|
$storage_1$ = $inst_1$.storage
$result_1$ = $storage_1$[$index_1$]
